---
title: "Marine Ecology Data Assignment"
author: "José Luis Acuña"
date: "1/8/2021"
output: html_document
---

```{r sets working directory}
setwd("D:/OneDrive - Universidad de Oviedo/activo/20210901-Marine Ecology/marine_ecology_data_assignment")

source("marecol_functions.R")

# plots 2D NMDS  with sample labels and colors for
# the cluster groups, such that. The arguments are: 
#     nmd   is the NMDS
#     clu   is the cluster
#     bran  is the number of branches (an integer > 1)
#     df    is the species dataframe

plot.NMDS <- function(nmd, clu, bran, df)

# shortens species names in file species.names.csv with first column as
# "Chirodota laevis [%] (C. laevis) * PI: Molis"
# Arguments are:
#     n1:      number of characters for the short genus name
#     n2:      number of characters for the short species name
#     sep:     separator character

shortens.names <- function(n1, n2, sep)


```

**PANGAEA.846449-tallus_disturbance_experiment**
https://doi.pangaea.de/10.1594/PANGAEA.846449

The effects of biotic disturbances, like seaweed whiplash, on the diversity of benthic communities are well documented for temperate coastal systems, yet missing for Arctic benthos. In Arctic soft-bottom habitats, kelp thalli occur either continuously (e.g. trapped by sediment) or sporadically (by drifting on the sediment) after detachment from rocky shores. To explore whether a kelp thallus can disturb the structure and diversity of a coastal Arctic soft-bottom assemblage, we continuously fixed a single thallus of the kelp Saccharina latissima to or sporadically (i.e. biweekly) moved it on the sediment and compared treatment effects to unmanipulated plots (= controls). On 6 September 2013 (i.e. after 73 days of manipulation), one sediment core was taken from each of the 30 plots (n = 10), from which the number of individuals of each of the 45 encountered animal species were recorded. Coded with a number (= block) and letter (= treatment, where capitol and minor letters indicate two experimental units from the same block; c = control, s = sporadic, f = fixed, i.e. continuous)

Use this code to open the species data:

tallus.disturbance <- read.csv("PANGAEA.846449-tallus_disturbance_experiment.csv",
  sep=";",          # separated by semicolon
  header = TRUE,    # reads column names from first row
  row.names = 1)    # reads row names from first columns


```{r tallus_disturbance_experiment}

# loads packages

library(tidyverse)
library(vegan)  


# opens file
tallus.disturbance <- read.csv("PANGAEA.846449-tallus_disturbance_experiment.csv",
  sep=";",          # separated by semicolon
  header = TRUE,    # reads column names from first row
  row.names = 1)    # reads row names from first columns

# creates a semimatrix with the bray-curtis simmilarities among all possible
# pairs of samples:
tallus.d <- vegdist(tallus.disturbance, "bray") 

# performs cluster analysis using the average distance between groups
# of samples:

tallus.cluster <- hclust(tallus.d, method="average")

# plots the result of the analysis, which is a dendrogram of simmilarities
# between samples:

plot(tallus.cluster) 

# Performs NMDS  to produce a 2D representation

tallus.NMDS<-metaMDS(tallus.disturbance,k=2) 

# stress is just too high, it requires transformation. We'll first do square root
# transformation

tallus.disturbance<-sqrt(tallus.disturbance)

# Performs NMDS  to produce a 2D representation

tallus.NMDS<-metaMDS(tallus.disturbance,k=2) 

# stress is again too high. Performs NMDS again to produce a 3D representation

tallus.NMDS<-metaMDS(tallus.disturbance,k=3) 

# plots NMDS using function plot.NMDS, using as input the NMDS (nmd), the
# cluster(clu), the number of branches (bra, an integer >1) and the species
# dataframe (df)

plot.NMDS (tallus.NMDS, tallus.cluster, 3, tallus.disturbance)


shortens.names(3,3,"_")
```

  